### 2016-03-09 hans/bayes-tardis-luminosity-prediction.tex
(wip) remove m

\usepackage[a4paper, left=20mm, right=20mm, top=25mm, bottom=25mm]{geometry}
\usepackage[numbers]{natbib}
%\usepackage{hyperref}
\synctex=1

@@

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{160228}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}
\parindent=0mm
\def\pp{\hfill\par \vspace{\baselineskip}}
%% \def\pp{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm,amssymb,amsfonts,amsmath}  % bold math symbols, AMStex
\usepackage{graphicx}    % standard graphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\lleq}[1]{\label{#1} }
% TO REMOVE THE LABEL LETTERS FROM THE EQUATION DISPLAY, COMMENT OUT THIS LINE:
\renewcommand{\lleq}[1]{\label{#1} {\scriptstyle {\rm (#1)}} \hspace*{2ex} }
% \renewcommand{\labelenumi}{\arabic{section}.\arabic{subsection}.\arabic{enumi}}
% \renewcommand{\thesubsubsection}{\arabic{subsubsection}}

@@

\DeclareMathOperator{\GammaDist}{Gamma}
\newcommand{\npack}{n_p}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{L_{\rm max}}

@@

  \item Separately, Fred introduces the idea of combining elementary
    $b$-bins into compound bins with index $m$ as this improves the
    statistics.

  \item The simplification for large $N$ etc permits a Gaussian
    approximation, hence Laplace, hence the key equation (pmp) may be
    simplified to some approximation, making use of the moment
    generating function eg Eq (mgd) unnecessary.

  \item In principle, the above sounds like a very good development.
    There are some important caveats:
    \begin{enumerate}
    \item I see no detail on how to find $a(\nu,\bmphi)$. The first
      question would be if there is some numerical indication for its
      form. This seems to be equivalent to asking what the dependence
      of (say) the Gamma parameter $\alpha$ on $\nu$ is, and that is
      of course of great interest anyway.

    \item My second comment is that $a(\nu,\bmphi)$ is probably just
      your normal parametrisation which enters best-fit algorithms.
      If we are to treat it as a best-fit parametrisation with
      metaparameters, then that may just bring back all the
      complications which we thought had disappeared. This would have
      to be investigated. If on the other hand you also want to
      treat $a(\nu,\bmphi)$ as some kind of ``constant function'',
      then one would have to set out how stable this constant function
      would be against various underlying changes. In other words, 
      if calculation of $a()$ is not part of the general algorithm,
      how is it fixed? and can it be fixed for all cases?

    \item I think there may be a problem of combining the two proposed
      simplifications $L \to X$ and $m$ compound bins. Depending on
      the variability of the parametrisation as function of $\nu$, the
      shape of $p(\bmL\cond N,\bmphi)$ for single bins with large
      $\Lmax$ may differ significantly from the shape for many small
      bins combined into one using $m$. In other words, rebinning
      and rescaling may not be compatible.
    \end{enumerate}

@@

  \item We treat the results of a single TARDIS run as \textit{data}
    %%%%%
    \fred{If in the end we want to compare to actual data from a
      telescope, I'd prefer to not call the TARDIS output
      ``data''. Why don't we just call it ``TARDIS output''?} %
    \hans{I agree, ``data'' is awkward, and it is indeed useful to
      keep data and $D$ for later. ``Output'' would be ok, except that
      we should change the symbol $D$ to something else too. Symbol
      $O$ is not good. For the moment I would suggest $D_S$ (subscript
      simulation) or $D_T$ since we know that $T$ will later stand for
      temperature. Another useful term would be ``simdata'', symbol
      $S$.}
    %%%%%
    $D$, i.e.\ fixed results which we do not question but accept as
    given.  This ``data'' consists of the total number of packets
    leaving the supernova $\npack$ plus the set of luminosities
    $\bml$ and frequencies $\bmnu$ of each packet,
  \begin{align}
    \lleq{dfc}
    D = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  In terms of the $\mathbf{1}_V(x)$ the indicator function requiring
  $x \in V$, the number of packets in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and the total luminosity in bin $b$ is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )\,,
  \end{align}
  with $\npack = \sum_b n_b$ and the total supernova luminosity $\ell
  = \sum_b \ell_b$.

\end{mtemize}
\hans{Discussion of rebinning with index $m$?}

@@

While we treat the TARDIS output as fixed ``data'' as discussed above,
a repeat of the simulation with a different Monte Carlo seed would
evidently yield different ``data'' $D'$ even when all other simulation
parameters remain the same. Clearly, $n_b$ is a specific realization
of a stochastic variable $N_b$ governed by a model pdf $p(N_b)$. %
\fred{In our general case, $\npack = \sum_b n_b$ is an instance of a
  random variable because some packets leave the supernova, some
  don't. I guess in most applications the total number of events would
  be chosen and not random. We are fortunate that this difference does
  not matter because we consider bins independently. } %
\hans{I will come back to this later} %
Similarly, $\bml_b$ is a specific realization of a collection of
stochastic variables $\bmL_b =\{L_{b,1},\ldots,L_{b,\smN_b}\}$, where
the number $N_b$ of such variables is itself stochastic. The relevant
pdf for $\bmL_b$ therefore always presupposes knowledge of $N_b$, so
we must work with the conditional probability $p(\bmL_b\cond
N_b)$. This situation is captured mathematically as $N_b \sim p(N_b),
\bmL_b \sim p(\bmL_b\cond N_b)$ where $\sim$ means ``is by hypothesis
distributed in this model according to the pdf'', i.e.\
\begin{align*}
  (N_b,\bmL_b) &\ \sim\ p(N_b,\bmL_b\cond \bmphi_b,\lambda_b)
  = p(\bmL_b\cond N_b,\bmphi_b)\; p(N_b\cond \lambda_b),
\end{align*}
where $\bmphi_b$ is the set of parameters of the model pdf which
govern $\bmL_b$ and $\lambda_b$ is the parameter of the model pdf
governing $N_b$.  To simulate $(N_b,\bmL_b)$, we clearly first need to
generate one $N_b$, followed by the generation of $N$ packet
luminosities $\bmL_b$. This situation is termed a \textit{compound
  process}.

@@

\item Based on the assumption of bin-bin statistical independence, we
  shall henceforth drop the bin index and do all our calculations for
  an individual bin $b$ only, writing for example $L_b \to L, n_b \to
  n$ etc.  The total spectrum is then obtained by doing any given
  analysis or calculation for all bins. But note that while the
  single-bin probability $p(N_b,\bmL_b\cond n_b, \bml,
  \bmnu,\bmphi,\bmlambda,\hmod)$ depends only on the number of packets
  $n_b$ in the bin, it depends on the luminosities and frequencies of
  \emph{all} packets.

@@

  \item In general, the distribution of $L_j$ could be a of a
  different functional form in each bin. For simplicity, we assume it
  is identical but may depend on parameters that differ from bin to
  bin. In practice, the distribution of packets does not change
  abruptly with frequency. We therefore assume that any parameter
  needed to fully specify $p(L_j)$ is given as a continuous function
  of frequency and parameters $\bmphi$; i.e.,
  \begin{align}
    \lleq{pmdb}
    p(L_j) = p(L_j \cond \nu, \bmphi)
  \end{align}
  For example, assume $p(L_j)$ depends on one parameter $a$. If we fit
  separately in each bin, we would have $B+1$ fits and $B+1$
  parameters. In each such fit, there would be only few samples to
  constrain $a$ in each bin. Then instead of fitting $a$ separately in
  each bin, we assume $a = a(\nu, \bmphi)$. This greatly reduces the
  number of parameters if $B$ is large because a handful of
  parameters, $\bmphi$, can be inferred from the complete set of
  packets. Necessarily, $a$ is best determined for the frequency
  around which the density of packets is highest. %
  \hans{I would replace the word ``sample'' above and throughout the
    text by ``packet''. The word ``sample'' refers in statistics to
    the entire data set $D$, not just to one data point.} %

@@

\item Priors: metaparameter $a$ \fred{?}  %
  \hans{This is a placeholder to remind us that metaparameters need to
    be discussed at some point. Perhaps not here, but we should
    remember.}
\end{mtemize}

@@

We first show how to use Bayesian prediction to find the probability
distribution for $N$, the number of packets in a bin, given knowledge
of ``data'' in the form of TARDIS multiplicity $n$ for that bin.
Since $n$ is a specific
realization of $N$, it results from the same Poisson distribution
\begin{align}
  \lleq{pmcb}
   p(n\cond\lambda) &= \frac{e^{-\lambda}\lambda^n}{n!}
\end{align}
with the same value for $\lambda$ as in (\ref{pmc}).  We can hence use
Bayes' Theorem to find the posterior distribution for $\lambda$ given
data $n$ and the Poisson hypothesis
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int d\lambda\, p(n\cond\lambda)\,p(\lambda)}
\end{align}
to find the prediction of $N$ given $n$ taking into account all
possible values of $\lambda$,
\begin{align}
  \lleq{pme}
  p(N\cond n)
  &= \int d\lambda\,p(N,\lambda\cond n)
  \ =\ \int d\lambda\,p(N\cond\lambda, n)\,p(\lambda\cond n)
  \ =\ \frac{\int d\lambda\,p(N\cond\lambda)\,p(n\cond\lambda)\,p(\lambda)}
  {\int d\lambda\, p(n\cond\lambda)\,p(\lambda)}
\end{align}\checked{}
where $p(N\cond\lambda,n) = p(N\cond\lambda)$ is independent of $n$.
Taking the prior (see Appendix)\footnote{For a uniform prior,
  the case $a{=}0$, we require of course a maximum $\lambda_{\rm
    max}$. As long as $\lambda_{\rm max}$ is large enough, the
  integral can effectively be taken to infinity.}
\begin{align}
  \lleq{pmf}
  p(\lambda) = c\lambda^{-a},
\end{align}
common choices are $a=0$ (uniform prior) and $a=1/2$ (Jeffreys
prior). Both choices lead to an improper prior but the posterior
predictive (see below) is proper.
Inserting \refeq{pmf} and (\ref{pmcb}) into (\ref{pme}), we obtain for
the evidence and the posterior predictive
\begin{align}
  \lleq{pmg}
  p(n) &= %
  \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
  \ =\
  c \int_0^\infty d\lambda\, \frac{e^{-\lambda}\,\lambda^{n-a}}{n!}
  \ =\ c\frac{(n-a)!}{n!}
  \\
  \lleq{pmh}
  p(N\cond n) %
  &= \frac{n!}{(n-a)!} \int_0^\infty d\lambda\,
  \frac{e^{-2\lambda} \lambda^{N+n-a}}{N!\, n!}
  \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}}
\end{align}
which we recognise as a Negative Binomial distribution describing the
probability to observe $N$ successes given $k=n+1-a$ failures in a
sequence of Bernoulli trials with success probability $\rho=1/2$
\begin{align}
  \lleq{pmhb}
  p(N\cond n) &= \NBD(N\cond k{=}n{+}1{-}a, \rho{=}\tfrac{1}{2})
  \ =\ \binom{N{+}n{-}a}{N}
  \left(\frac{1}{2}\right)^N \left(\frac{1}{2}\right)^{n+1-a}
\end{align}
We shall use the properties of the Negative Binomial in the
calculations below. For non-integer $a$, the Negative Binomial
generalizes immediately to the Pólya distribution by replacing $x! \to
\Gamma(x+1)$. Note that the distribution of $N$ given $n$ is
independent of $\bmL$ and $\bml$ by construction.

@@

@@

\fred{We should be more general: let's reserve $n$ for the number of
  packets observed in a bin. Suppose this is $n=1$. From that, we can
  hardly infer the luminosity distribution. Instead, we group the
  packets in larger bins such that there are at least $m \gtrsim 300$
  packets, and then assume the distribution doesn't change much with
  frequency.

  Hence we will learn $\bmphi$ from $m$ samples, and
  $p(\bmphi\cond N,n,\bml) \to p(\bmphi\cond N,m,\bml)$ and we can
  keep $\bml$ with the same meaning. So we replace $n \to m$ below
  except in $p(N \cond n$)}

Our goal of predicting $L$ given the TARDIS output $(\npack,\bml,
\bmnu)$ is captured in the probability $p(L\cond \npack,\bml, \bmnu)$
with all relevant parameters integrated out. We first note from
\refeq{pmh} that $p(N\cond \npack,\bml, \bmnu) = p(N\cond n)$ does not
depend on $\bml$ or $\bmnu$ directly but only through $n$, the number
of packets in the bin; cf. \refeq{dff}. Furthermore, since $L$ is
fully determined by $(N,\bmL)$,
\begin{align}
  \lleq{pmjd}
  L &= \sum_{j=1}^N L_j
\end{align}
is does not depend directly on $(\npack),\bml,\bmnu)$, i.e.\
\begin{align}
  \lleq{pmi}
  p(L\cond N,\bmL,n,\bml, \bmnu)
  = p(L\cond N,\bmL)
  &= \delta(L - \textstyle\sum_{j=1}^N L_j).
\end{align}
Hence the prediction is given by
\begin{align}
  \lleq{pmj}
  p(L\cond n,\bml, \bmnu)
  &= \sum_{N=0}^\infty p(L\cond N,n,\bml, \bmnu)\,p(N\cond n,\bml, \bmnu) \\
  \lleq{pmjb}
  &= \sum_N p(N\cond n)\,\int d\bmL\, p(L\cond N,\bmL,n,\bml, \bmnu)\, p(\bmL\cond N,n,\bml, \bmnu)
  \\
  \lleq{pmjc}
  &= \sum_N p(N\cond n)\,\int d\bmL\,\delta(L - \textstyle\sum_{j=1}^N L_j)
  \, p(\bmL\cond N,n,\bml, \bmnu)
\end{align}
\fred{Add missing $\bmnu$! Everything that is conditional on $\bml$ is
  also conditional on $\bmnu$} The dependence on the TARDIS ``data''
$(\npack,\bml, \bmnu)$ enters through $p(\bmL\cond N,\npack,\bml,
\bmnu)$ in terms of the parameters $\bmphi$ of $p(\bmL\cond
N,\bmphi)$,
\begin{align}
  \lleq{pmk}
  p(\bmL\cond N,\npack,\bml, \bmnu)
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi,\npack,\bml, \bmnu)\,p(\bmphi\cond N,\npack,\bml, \bmnu) \\
  &= \int d\bmphi\, p(\bmL\cond N,\bmphi)\,p(\bmphi\cond N,\npack,\bml, \bmnu).
\end{align}
because the likelihood for $\bmL$ is fully determined by $N,\bmphi$
and hence independent of $(\npack,\bml, \bmnu)$.
%
Again using Bayes' Theorem (\ref{bth})
\begin{align}
  \lleq{pmkb}
  p(\bmphi\cond N,\npack,\bml, \bmnu) &=
  \frac{p(\bml\cond N,\npack,\bmphi,\bmnu)\,p(\bmphi\cond N,\npack,\bmnu)} %
  {\int d\bmphi\,p(\bml\cond N,\npack,\bmphi,\bmnu)\,p(\bmphi\cond N,\npack,\bmnu)}, %
  % \ =\ \frac{p(\bml\cond N,\npack,\bmphi)\,p(\bmphi\cond N,n)} %
  % {p(\bml\cond n)} %
\end{align}

>>>>

\usepackage[a4paper, left=20mm, right=20mm, top=25mm, bottom=25mm]{geometry}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\synctex=1

@@

%% CHANGE THIS DATE AS YOU MODIFY THE FILE
\newcommand{\vdate}{160309}
\pagestyle{myheadings}
\markboth{Version \vdate} {Version \vdate}
\parindent=0mm
\def\pp{\hfill\par \vspace{\baselineskip}}
%% \def\pp{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{bm,amssymb,amsfonts,amsmath}  % bold math symbols, AMStex
\usepackage{graphicx}    % standard graphics
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\lleq}[1]{\label{#1} }
% TO REMOVE THE LABEL LETTERS FROM THE EQUATION DISPLAY, COMMENT OUT THIS LINE:
\renewcommand{\lleq}[1]{\label{#1} {\scriptstyle {\rm (#1)}} \hspace*{2ex} }
% \renewcommand{\labelenumi}{\arabic{section}.\arabic{subsection}.\arabic{enumi}}
% \renewcommand{\thesubsubsection}{\arabic{subsubsection}}

@@

\DeclareMathOperator{\GammaDist}{Gamma}
\newcommand{\npack}{n_p}
\newcommand{\lmax}{\ell_{\rm max}}
\newcommand{\Lmax}{L_{\rm max}}
\newcommand{\rmdx}[1]{\mbox{d} #1 \,} % differential

@@

    \item Separately, Fred introduces the idea of combining elementary
    $b$-bins into compound bins with index $m$ as this improves the
    statistics.%
    \fred{This was an intermediate idea, I now removed any reference
      to it. There should be no more binning when infering $\bmphi$}

    \item The simplification for large $N$ etc permits a Gaussian
    approximation, hence Laplace, hence the key equation (pmp) may be
    simplified to some approximation, making use of the moment
    generating function eg Eq (mgd) unnecessary.%
    \fred{Yes, so you put in a lot of work to derive the equations and
      I checked them carefully but I think in the end we will not need
      them and we should then remove them from the draft}

  \item In principle, the above sounds like a very good development.
    There are some important caveats:
    \begin{enumerate}
      \item I see no detail on how to find $a(\nu,\bmphi)$. The first
      question would be if there is some numerical indication for its
      form. This seems to be equivalent to asking what the dependence
      of (say) the Gamma parameter $\alpha$ on $\nu$ is, and that is
      of course of great interest anyway.%
      \fred{Yes, that's equivalent. I just started with a polynomial
        for $a$ but that's ad-hoc.}

      \item My second comment is that $a(\nu,\bmphi)$ is probably just
      your normal parametrisation which enters best-fit algorithms.
      If we are to treat it as a best-fit parametrisation with
      metaparameters, then that may just bring back all the
      complications which we thought had disappeared. This would have
      to be investigated. If on the other hand you also want to treat
      $a(\nu,\bmphi)$ as some kind of ``constant function'', then one
      would have to set out how stable this constant function would be
      against various underlying changes. In other words, if
      calculation of $a()$ is not part of the general algorithm, how
      is it fixed? and can it be fixed for all cases?%
      \fred{I don't see what you mean by ``constant function''. The
        choice of $a$ and how it depends on $\bmphi$ is part of the
        overal model.}

      \item I think there may be a problem of combining the two
      proposed simplifications $L \to X$ and $m$ compound
      bins. Depending on the variability of the parametrisation as
      function of $\nu$, the shape of $p(\bmL\cond N,\bmphi)$ for
      single bins with large $\Lmax$ may differ significantly from the
      shape for many small bins combined into one using $m$. In other
      words, rebinning and rescaling may not be compatible.%
      \fred{No more compound bins, we just do an unbinned fit wrt
        $\bmphi$, so there is no issue}
    \end{enumerate}

@@

  \item We treat the results of a single TARDIS run as \textit{simulated data}
    %%%%%
    \fred{If in the end we want to compare to actual data from a
      telescope, I'd prefer to not call the TARDIS output
      ``data''. Why don't we just call it ``TARDIS output''?} %
    \hans{I agree, ``data'' is awkward, and it is indeed useful to
      keep data and $D$ for later. ``Output'' would be ok, except that
      we should change the symbol $D$ to something else too. Symbol
      $O$ is not good. For the moment I would suggest $D_S$ (subscript
      simulation) or $D_T$ since we know that $T$ will later stand for
      temperature. Another useful term would be ``simdata'', symbol
      $S$.}
    %%%%%
    $S$, i.e.\ fixed results which we do not question but accept as
    given.  These ``data'' consist of the total number of packets
    leaving the supernova $\npack$ plus the set of luminosities
    $\bml$ and frequencies $\bmnu$ of each packet,
  \begin{align}
    \lleq{dfc}
    S = \{ \npack, \bml, \bmnu \} \,,
  \end{align}
  where
  \begin{align}
    \lleq{dfd}
    % \bml_b &= (\ell_{b,1},\ell_{b,2},\ldots,\ell_{b,n_b})
    \dim \bml = \dim \bmnu = \npack \,.
  \end{align}
  In terms of $\mathbf{1}_V(x)$---the indicator function requiring
  $x \in V$---the number of packets in bin $b$ is
  \begin{align}
    \lleq{dff}
    n_b = \sum_{j=1}^{\npack} \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )
  \end{align}
  and the total luminosity in bin $b$ is
  \begin{align}
    \lleq{dfe}
    \ell_b = \sum_{j=1}^{\npack}\ell_j \mathbf{1}_{[\nu_b,\nu_{b+1}]}( \nu_j )\,,
  \end{align}
  with $\npack = \sum_b n_b$ and the total supernova luminosity $\ell
  = \sum_b \ell_b$.

@@

While we treat the TARDIS output as fixed ``data'' as discussed above,
a repeat of the simulation with a different Monte Carlo seed would
evidently yield different ``data'' $S'$ even when all other simulation
parameters remain the same. Clearly, $n_b$ is a specific realization
of a stochastic variable $N_b$ governed by a model pdf $p(N_b)$. %
\fred{In our general case, $\npack = \sum_b n_b$ is an instance of a
  random variable because some packets leave the supernova, some
  don't. I guess in most applications the total number of events would
  be chosen and not random. We are fortunate that this difference does
  not matter because we consider bins independently. } %
\hans{I will come back to this later} %
Similarly, $\bml_b$ is a specific realization of a collection of
stochastic variables $\bmL_b =\{L_{b,1},\ldots,L_{b,\smN_b}\}$, where
the number $N_b$ of such variables is itself stochastic. The relevant
pdf for $\bmL_b$ therefore always presupposes knowledge of $N_b$, so
we must work with the conditional probability $p(\bmL_b\cond
N_b)$. This situation is captured mathematically as $N_b \sim p(N_b),
\bmL_b \sim p(\bmL_b\cond N_b)$ where $\sim$ means ``is by hypothesis
distributed in this model according to the pdf'', i.e.\
\begin{align*}
  (N_b,\bmL_b) &\ \sim\ p(N_b,\bmL_b\cond \bmphi_b,\lambda_b)
  = p(\bmL_b\cond N_b,\bmphi_b)\; p(N_b\cond \lambda_b),
\end{align*}
where $\bmphi_b$ is the set of parameters of the model pdf which
govern $\bmL_b$ and $\lambda_b$ is the parameter of the model pdf
governing $N_b$.  To simulate $(N_b,\bmL_b)$, we clearly first need to
generate one $N_b$, followed by the generation of $N$ packet
luminosities $\bmL_b$. This situation is termed a \textit{compound
  process}.

@@

\item Based on the assumption of bin-bin statistical independence, we
  shall henceforth drop the bin index and do all our calculations for
  an individual bin $b$ only, writing for example $L_b \to L, n_b \to
  n$ etc.  The total spectrum is then obtained by doing any given
  analysis or calculation for all bins. But note that while the
  single-bin probability $p(N_b,\bmL_b\cond n_b, \bml,
  \bmnu,\bmphi,\bmlambda,\hmod)$ depends only on the number of packets
  $n_b$ in the bin, it depends on the luminosities and frequencies of
  \emph{all} packets.
  %
  \fred{Perhaps it doesn't simplify the reading experience to omit $b$?}

@@

  \item In general, the distribution of $L_j$ could be a of a
  different functional form in each bin. For simplicity, we assume it
  is identical but may depend on parameters that differ from bin to
  bin. In practice, the distribution of packets does not change
  abruptly with frequency. We therefore assume that any parameter
  needed to fully specify $p(L_j)$ is given as a continuous function
  of frequency and parameters $\bmphi$; i.e.,
  \begin{align}
    \lleq{pmdb}
    p(L_j) = p(L_j \cond \nu, \bmphi)
  \end{align}
  For example, assume $p(L_j)$ depends on one parameter $a$. If we fit
  separately in each bin, we would have $B+1$ fits and $B+1$ copies of
  $a$. In each fit, there would be only few packets to constrain
  $a$. To avoid that, we assume $a = a(\nu, \bmphi)$. For concreteness,
  assume $\bmphi = (a_0, a_1)$ and
  \begin{align}
    \lleq{pmdc}
    a(\nu, \bmphi) = a_0 +a_1 \nu
  \end{align}
  This greatly reduces the number of parameters from $B+1$ to 2 and
  ensures that $a$ is continuous in $\nu$. $a$, or equivalently
  $\bmphi$, is now determined from a single fit to all packets. The
  uncertainty on $a$ is at a minimum for the frequency around which
  the density of packets is highest. \fred{Refer to plot with
    uncertainty band}

@@

  \item Priors: metaparameter $a$
  %
  \hans{This is a placeholder to remind us that metaparameters need to
    be discussed at some point. Perhaps not here, but we should
    remember.}
\end{mtemize}

@@

We first show how to use Bayesian prediction to find the probability
distribution for $N$, the number of packets in a bin, given knowledge
of ``data'' in the form of TARDIS multiplicity $n$ for that bin.
Since $n$ is a specific
realization of $N$, it results from the same Poisson distribution
\begin{align}
  \lleq{pmcb}
   p(n\cond\lambda) &= \frac{e^{-\lambda}\lambda^n}{n!}
\end{align}
with the same value for $\lambda$ as in (\ref{pmc}).  We can hence use
Bayes' Theorem to find the posterior distribution for $\lambda$ given
data $n$ and the Poisson hypothesis
\begin{align}
  \lleq{pmd}
  p(\lambda\cond n) %
  &= \frac{p(n\cond\lambda)\,p(\lambda)} {p(n)}
  \ =\ \frac{p(n\cond\lambda)\,p(\lambda)} %
  {\int d\lambda\, p(n\cond\lambda)\,p(\lambda)}
\end{align}
to find the prediction of $N$ given $n$ taking into account all
possible values of $\lambda$,
\begin{align}
  \lleq{pme}
  p(N\cond n)
  &= \int d\lambda\,p(N,\lambda\cond n)
  \ =\ \int d\lambda\,p(N\cond\lambda, n)\,p(\lambda\cond n)
  \ =\ \frac{\int d\lambda\,p(N\cond\lambda)\,p(n\cond\lambda)\,p(\lambda)}
  {\int d\lambda\, p(n\cond\lambda)\,p(\lambda)}
\end{align}
where $p(N\cond\lambda,n) = p(N\cond\lambda)$ is independent of $n$.
Taking the prior (see Appendix)\footnote{For a uniform prior,
  the case $a{=}0$, we require of course a maximum $\lambda_{\rm
    max}$. As long as $\lambda_{\rm max}$ is large enough, the
  integral can effectively be taken to infinity.}
\begin{align}
  \lleq{pmf}
  p(\lambda) = c\lambda^{-a},
\end{align}
common choices are $a=0$ (uniform prior) and $a=1/2$ (Jeffreys
prior). Both choices lead to an improper prior but the posterior
predictive (see below) is proper.
Inserting \refeq{pmf} and (\ref{pmcb}) into (\ref{pme}), we obtain for
the evidence and the posterior predictive
\begin{align}
  \lleq{pmg}
  p(n) &= %
  \int_0^\infty d\lambda\, p(n\cond\lambda)\,p(\lambda)
  \ =\
  c \int_0^\infty d\lambda\, \frac{e^{-\lambda}\,\lambda^{n-a}}{n!}
  \ =\ c\frac{(n-a)!}{n!}
  \\
  \lleq{pmh}
  p(N\cond n) %
  &= \frac{n!}{(n-a)!} \int_0^\infty d\lambda\,
  \frac{e^{-2\lambda} \lambda^{N+n-a}}{N!\, n!}
  \ =\ \frac{(N+n-a)!}{N!\,(n-a)!\,2^{N+n+1-a}}
\end{align}
which we recognise as a Negative Binomial distribution describing the
probability to observe $N$ successes given $k=n+1-a$ failures in a
sequence of Bernoulli trials with success probability $\rho=1/2$
\begin{align}
  \lleq{pmhb}
  p(N\cond n) &= \NBD(N\cond k{=}n{+}1{-}a, \rho{=}\tfrac{1}{2})
  \ =\ \binom{N{+}n{-}a}{N}
  \left(\frac{1}{2}\right)^N \left(\frac{1}{2}\right)^{n+1-a}
\end{align}
We shall use the properties of the Negative Binomial in the
calculations below. \fred{Check if we still do later} For non-integer $a$,
the Negative Binomial generalizes immediately to the Pólya
distribution by replacing $x! \to \Gamma(x+1)$. Note that the
distribution of $N$ given $n$ is independent of $\bmL$ and $\bml$ by
construction but depends on $\bmnu$ implicitly through $n$.

@@

\fred{One thing is bugging me: To predict the luminosity in a bin, I
  should do this $ \int_{\nu_b}^{\nu_{b+1}} d{\nu} \, p(L, \nu) =
  \int_{\nu_b}^{\nu_{b+1}} d{\nu} \, p(L \cond \nu) p(\nu) $. We want
  to avoid modeling $p(\nu)$ because of the absorption lines, so we
  assume $p(L \cond \nu)$ approx. indep. of $\nu$ in the bin, so $ \approx p(L \cond
  \nu ) \int_{\nu_b}^{\nu_{b+1}} d{\nu} \, p( \nu) \equiv p(L \cond
  \nu) \, p_b$. Since $p_b$ is unknown, we should marginalize over it
  using our knowledge from Tardis output:
  $$
  p(L \cond \mbox{bin } b, \npack, \bmnu, \bml) = \int d{p_b}\, p(L \cond \nu) \, p_b \, p(p_b \cond \npack, \bmnu, \bml)
  $$

  So how can we turn $\int \rmdx{p_b} p_b \to \sum_N$? Where does
  $\npack$ come into play? If we can answer this, then we have one
  more hidden assumption clarified in the derivation below that was
  written for a binned analysis but now we want to be unbinned wrt
  $\bmnu, \bml$.  }

@@

% \fred{We should be more general: let's reserve $n$ for the number of
%   packets observed in a bin. Suppose this is $n=1$. From that, we can
%   hardly infer the luminosity distribution. Instead, we group the
%   packets in larger bins such that there are at least $m \gtrsim 300$
%   packets, and then assume the distribution doesn't change much with
%   frequency.

%   Hence we will learn $\bmphi$ from $m$ samples, and
%   $p(\bmphi\cond N,n,\bml) \to p(\bmphi\cond N,m,\bml)$ and we can
%   keep $\bml$ with the same meaning. So we replace $n \to m$ below
%   except in $p(N \cond n$)}

Our goal of predicting $L$ in bin $b$ given the TARDIS output
$(\npack,\bml, \bmnu)$ is captured in the probability $p(L\cond \nu,
\npack,\bml, \bmnu)$ with all relevant parameters integrated out using
the law of total probability.

To simplify the treatment, we assume that the frequency bins are
narrow or equivalently that $p(L \cond \nu, \npack, \bml, \bmnu)$ does
not vary within a bin. This implies that all hypothetical samples
$L_j$ in the bin follow the same distribution.  As the reference
frequency, we choose the bin center;
% and avoid integration over the bin
i.e., for each bin $b$, we set
\begin{align}
  \lleq{pmq}
  p(L\cond \nu, \npack,\bml, \bmnu) = p(L\cond (\nu_b + \nu_{b+1}) / 2, \npack,\bml, \bmnu)  \; \forall \nu \in [\nu_b, \nu_{b+1}] \, .
\end{align}
